\myfont Implementation and evaluation of register allocation for translating LLVM-{}- to x86-64

### 2024-01-14 article.tex
move factor

The instruction set used in this paper will be a union of the sets used in the 2022 and 2023 compilers courses in order to work as a drop-in replacement of LLVM for either of the two respective source languages:  Tiger and Dolphin. This instruction set is a subset of the one used in practice, as, for instance, neither of the languages implemented support interrupts, floating point operations etc., and instead only strive to cover the basics of compilers. In additional some other instructions are also included, like \lstinline!trunc! and \lstinline!switch!, to widen the range of supported frontend languages with the clang compiler.
%been added in order to help cover more generated LLVM. % FIXME
%with only four distinct block terminators 
%\begin{enumerate}
%    \item Unconditional \lstinline!br! followed by the label to branch to
%    \item Conditional \lstinline!br!  followed by an \lstinline!i1! condition and two labels
%    \item \lstinline!ret! followed by an optional return value
%    \item \lstinline!unreachable!
%\end{enumerate}
%The branching for most of these is trivial: after successful execution the flow of all but \lstinline!br!, \lstinline!ret! and \lstinline!unreachable! will unconditionally attempt to execute the next instruction in memory (i.e. increment the instruction pointer by instruction length).

@@

\noindent The measures in Table \ref{tab:dumb-fib}  are significantly equivalent to one another, with all allocators (except greedy 0) seeming to converge towards being 37\% slower than Clang. This isn't particularly surprising, as the x86 translation is more concerned with preserving callee saved registers than necessary in pushing each and every one despite not being in use. Also, instead of pushing with the \lstinline!push! instruction it would be faster to simply subtract from the \lstinline!%rsp! register directly and using \lstinline!mov! directly.

@@

@@

\caption{\label{tab:factori32}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}

>>>>

The instruction set used in this paper will be a combination of those used in the 2022 and 2023 compilers courses in order to work as a drop-in replacement of LLVM for either of the two respective source languages:  Tiger and Dolphin. This instruction set is a subset of the one used in practice, as, for instance, neither of the languages implemented support interrupts, floating point operations etc., and instead only strive to cover the basics of compilers. In additional some other instructions are also included, like \lstinline!trunc! and \lstinline!switch!, to widen the range of supported frontend languages with the clang compiler.
%been added in order to help cover more generated LLVM. % FIXME
%with only four distinct block terminators 
%\begin{enumerate}
%    \item Unconditional \lstinline!br! followed by the label to branch to
%    \item Conditional \lstinline!br!  followed by an \lstinline!i1! condition and two labels
%    \item \lstinline!ret! followed by an optional return value
%    \item \lstinline!unreachable!
%\end{enumerate}
%The branching for most of these is trivial: after successful execution the flow of all but \lstinline!br!, \lstinline!ret! and \lstinline!unreachable! will unconditionally attempt to execute the next instruction in memory (i.e. increment the instruction pointer by instruction length).

@@

\noindent The measures in Table \ref{tab:dumb-fib}  are remarkably similar to one another, with all allocators (except greedy 0) seeming to converge towards being 37\% slower than Clang. This isn't particularly surprising, as the x86 translation is more concerned with preserving callee saved registers than necessary in pushing each and every one despite not being in use. Also, instead of pushing with the \lstinline!push! instruction it would be faster to simply subtract from the \lstinline!%rsp! register directly and using \lstinline!mov! directly.

@@

\caption{\label{tab:factori32}Benchmark of \texttt{benches/factori32.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1000}}

@@
